---
title: "Introduction to BMGM"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to BMGM}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

The **BMGM** package implements a Bayesian pairwise graphical model for estimating
conditional independencies between variables of mixed types. It supports
continuous, discrete (count), zero-inflated count, and categorical data within a
single unified framework. The model uses a spike-and-slab prior for graph
estimation and an MCMC algorithm for posterior inference. It can also handle
missing data through built-in imputation during the estimation process.

For full methodological details, see Florez, M., Gottard, A., McAdams, C.,
Guindani, M., & Vannucci, M. (2025). A Bayesian approach for inference on mixed
graphical models. *Bayesian Analysis*, Advance Publication, 1-28.
[https://doi.org/10.1214/25-BA1557](https://doi.org/10.1214/25-BA1557)

## Getting started

```{r setup}
library(BMGM)
```

### Simulating mixed data

We begin by generating a simple dataset with four variables of different types:
a discrete count variable (`X1`), two continuous variables (`X2`, `X3`) where
`X3` depends on `X2`, and an independent categorical variable (`X4`) with four
categories.

```{r simulate}
set.seed(123)
n <- 200

X1 <- rpois(n, lambda = 3)
X2 <- rnorm(n, mean = 5)
X3 <- rnorm(n, mean = X2)
X4 <- sample(1:4, size = n, replace = TRUE)

X <- cbind(X1, X2, X3, X4)
```

### Specifying variable types

The `type` argument tells `bmgm()` how to treat each column. The options are:

- `"c"` — continuous (Gaussian)
- `"d"` — discrete (count)
- `"z"` — zero-inflated count
- `"m"` — categorical (multinomial)

```{r types}
type <- c("d", "c", "c", "m")
```

### Fitting the model

We fit the Bayesian Mixed Graphical Model using `bmgm()`. The `nburn` argument
controls the number of burn-in iterations, and `nsample` controls the number of
posterior samples retained for inference.

```{r fit, results='hide'}
fit <- bmgm(X, type, nburn = 5000, nsample = 10000)
```

### Inspecting results

Printing the fitted object gives a quick summary of the estimated graph:

```{r print}
fit
```

The model correctly identifies the edge between `X2` and `X3`, reflecting the
dependence we built into the simulation. The independent variables (`X1` and
`X4`) show no edges, as expected.

### Adjacency matrix and edge weights

The estimated adjacency matrix and edge weights are available as components of
the fitted object:

```{r adjacency}
fit$adj_G
```

```{r weights}
round(fit$adj_Beta, 3)
```

### Visualizing the graph

The estimated graph can be visualized with the **qgraph** package:

```{r plot, eval=FALSE}
qgraph::qgraph(fit$adj_G, labels = colnames(X))
```

## Handling missing data

BMGM can handle data with values missing at random. Simply include `NA` values
in the data matrix:

```{r missing, eval=FALSE}
X_miss <- X
X_miss[sample(1:n, 10), 2] <- NA
X_miss[sample(1:n, 10), 3] <- NA

fit_miss <- bmgm(X_miss, type, nburn = 5000, nsample = 10000)
fit_miss
```

## Further reading

- Florez, M., Gottard, A., McAdams, C., Guindani, M., & Vannucci, M. (2025).
  A Bayesian approach for inference on mixed graphical models. *Bayesian
  Analysis*, Advance Publication, 1-28.
  [https://doi.org/10.1214/25-BA1557](https://doi.org/10.1214/25-BA1557)
- Package source: <https://github.com/mauroflorez/BMGM>
